Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2020-10-27 paper/main.tex
paper

@@

@@

The exact MLE for $\sigma_0$ is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}. Unfortunately,  obtaining such ranking is NP-hard~\cite{DwoKumNao2001rank}. An alternative to the Kemeny ranking is the Borda ranking~\cite{AliMei2011kemeny}, which has several advantages: (i) it requires polynomial computational time; (ii) it  guaranties  high quality parameters as an estimator for the MM~\cite{CarProSha2013votes}; and (iii) in a general optimization perspective, Borda is an approximation to Kemeny~\cite{CopFleRur2010ordering}.\MANUEL{I'm not sure ``in a general optimization perspective'' means}

\subsection{Unbalanced Borda}\label{sec:uborda}
We introduce now the core for our probabilistic algorithm, the uBorda algorithm. Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking. The uBorda algorithm is a recent generalization of Borda~\cite{???}.\MANUEL{X This paper is missing \ekhine{}{to be pocessed by arxiv}} In uBorda, we consider that each permutation in the sample $\sigma\in S$ has an associated weight $w(\sigma)$ where $w\colon S_n \to \mathbb{R^{+}}$. Then, the score of \manuel{item}{position} $i$ is defined as follows:
%
\begin{equation}\label{eq:uborda_score}
B(i) =  \sum_{t\in S}  w(\sigma_t) \sigma_t(i) \enspace,
\end{equation}
%
and, the same as Borda, orders the \ekhine{items}{positions}\MANUEL{X why items? we talk about positions above. We could talk about decision variables if that is clearer} increasingly by their score.

@@

As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $\minit$ permutations that are uniformly randomly generated (although it would be possible to employ Latin Hyper-Cube sampling or other methods) and evaluated. Then, using the current sample $S$, the uBorda algorithm is applied to find a reference permutation $\hat{\sigma}_0$ (line~\ref{line:uborda}). Together with the parameter $\theta$, they define the Mallows model $MM(\hat\sigma_0, \theta)$. A decreasing $\theta$ parameter (line~\ref{line:theta}) implies a decreasing variance of the Mallows model (Eq.~\ref{eq:MM}), as explained in  Section~\ref{sec:sampling} below. Then, we sample a new candidate permutation from  $MM(\hat\sigma_0, \theta)$ (line~\ref{line:sample}). This candidate permutation (solution) is evaluated and added to the sample $S$. The process is repeated up to a given maximum budget of evaluations $\FEmax$.\MANUEL{If the learning step happens later, when does it happen here?}
\MANUEL{We do not explain the initial value of $\theta$ or how we decrease it. Can we give a bit more detail?}

@@

\manuel{At each iteration of the algorithm the following steps take place: (1) randomly
generate a permutation distributed according to the internal model and include
it in the sample and (2) evaluate the new permutation and update the
probabilistic model. These steps correspond to two statistical operations: (1)
sample a probability distribution and (2) learn a probability distribution of
the transformation of the sample. Both are describe in the subsequent sections.}{}\MANUEL{I don't think we need this, it is said above (if some detail is missing we should integrate it the paragraph above)}

@@

\subsubsection{Learning step.}
At each iteration we start by fitting a MM to the sample $S$, which implies
estimating $\hat\sigma_0$ and $\theta$. Parameter $\theta$ is updated
deterministically at the sampling step.  The parameter $\sigma_0$, however, is
learned from the current sample $S$ by using the uBorda algorithm with a weight
$w(\sigma)=\rho^{f(\sigma)}$ (Section~\ref{sec:uborda}). The learning rate is
controlled by parameter $\rho \in (0,1)$. %
In UMM, the value of $\rho$ depends on the set of true fitness evaluations $F$
and it is set at each iteration such that the largest $100r_1$\% of the mass of
the weights is concentrated in the best $100r_2$\% of the solutions in $S$, i.e.,
\MANUEL{Let's talk about this Equation tomorrow}
%
\begin{equation}
\begin{split}
r_1 \sum_{\sigma\in S}\rho^{f(\sigma)} =  \sum_{\sigma\in S'}\rho^{f(\sigma)} \quad\text{where}\quad |S'| = r_2|S| \land \forall \sigma'\in S', \sigma\in S\setminus S' : f(\sigma')\geq f(\sigma)
\end{split}
\end{equation}
%
\MANUEL{should the $\sigma$ be the same on both sides of $=$?}
where $S'$ contains the $r_2|S|$ elements of $S$ with the best corresponding fitness values in $F$. The values of $r_1, r_2 \in (0, 1)$ are user-defined and control the balance between exploration and intensification: higher values bias the sampling towards the best solutions found so far, which may lead to quick stagnation, whereas smaller values make the sampling more uniform, thus slowing down the learning.

@@

\subsubsection{Sampling step.}\label{sec:sampling}
Sampling from $MM(\sigma_0,\theta)$  can be done efficiently~\cite{IruCalLoz2016permallows} given the parameters  $\sigma_0$ and $\theta$, as studied by \citet{Collas,IruCalLoz2016permallows,DoiPekReg2004rank}\MANUEL{Which paper is Collas? It is not found.} where the computational complexities are $O(n\log n)$ and $O(n^2)$.\MANUEL{respectively?}
%
At each iteration of UMM, $\sigma_0=\hat\sigma_0$ is the permutation obtained in the previous learning step. The scale $\theta$ controls the variance and expected distance $\Edist$ of the MM, i.e., $D$ is the distance of a random permutation $\sigma\sim MM(\sigma_0, \theta)$ to the location parameter $\sigma_0$:\MANUEL{where does this formula come from?}
%
\begin{equation}\label{eq:expectation}
\Edist = \frac{n \cdot \exp(-\theta)}{1-\exp(-\theta)} - \sum_{j=1}^{n}  \frac{j\cdot\exp(-j  \theta)}{1-\exp(-j\theta)}
\end{equation}

@@

A purely explotation approach would always generate $\sigma_0$ in the sampling stage. However, the setting of a decreasing variance \MANUEL{But we are controlling the expectation not the variance???}  controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.  Similar approaches have been taken in different contexts~\cite{ArzCebPer2019qap}.

@@

\subsubsection{Other}.  For simplicity, we use Kendall's $\tau$
distance~(Eq.~\ref{eq:kendall_decomp}) in all
experiments\MANUEL{\citet{ZaeStoBar2014:ppsn} claims this is called Swap
  distance, could you double-check?}. Nevertheless,
\citet{ZaeStoFriFisNauBar2014} points out that CEGO performs best on the QAP
when using Hamming distance.  We plan to extend UMM to other distance measures
and to the dynamic selection of distance measures~\citep{ZaeStoBar2014:ppsn}.
In all experiments, we consider a maximum budget of $\FEmax=400$ evaluations of
the actual objective function. In a white-box context, state-of-the-art
algorithms for the problems considered here typically evaluate
thousands of solutions, thus, the budget considered
here for the black-box context is extremely limited.

@@

\subsubsection{Computing environment.}
All experiments were run on Intel Xeon ``Ivybridge'' E5-2650v2 CPUs at
2.60\,GHz, 64\,GB RAM running CentOS Linux release 7.4.1708 (Core).
\MANUEL{This may need updating since you ran some experiments? didn't you?}

@@

\ekhine{}{In prev sections we say that we show here how to tune $rho$. Le ves sitio? con los heatmaps? al principio de la sección? an outline here}
\ekhine{}{recall that $w(\sigma) = \rho^{f(sigma)}$. We select $\rho$ in such a way that the 100$r_1\%$ of the samples have the 100$w\%$ of the weight}

The first part of the experimental analysis concerns the \manuel{tunning}{analyses} of parameter $\rho$. As stated in Section~\ref{sec:sampling}, $\rho$ is automatically set at each iteration in such a way that the $100r_1\%$ of the samples have the $100r_2\%$ of the weight. In this regard, we tried different configurations of parameters $r_1$ and $r_2$ in all the instances.\MANUEL{why did we try this values.} Figure~\ref{fig:heatmaps} shows the best solution\MANUEL{best solution found at the end of a run?} (averaged for 10 runs) for instances: \texttt{N-t1d100.01} and \texttt{N-t1d150.01} for the LOP, \texttt{reC05} for the PFSP, \texttt{kra32} for the QAP. \MANUEL{The caption says synthetic instances}

\begin{figure}[htb]
  \centering%
  \includegraphics[width=0.24\textwidth]{../img/heatmap_N-t1d100_01}
  \includegraphics[width=0.24\textwidth]{../img/heatmap_N-t1d150_01}
  \includegraphics[width=0.24\textwidth]{../img/heatmap_rec05}
  \includegraphics[width=0.24\textwidth]{../img/heatmap_kra32_dat}
    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on LOP synthetic instances.\label{fig:heatmaps}}
\end{figure}

@@

Over all the instances \MANUEL{(which instances we tested? all synthetic?)} we
see that the configuration $r_1 =0.1$, $r_2=0.9$ has good
performance\MANUEL{how do you define this? The best mean?}. From this
preliminary experiment, it is not obvious how the best settings are related to
problem features, such as permutation length ($n$). On the other hand, given
the sharp transitions in the plots, it is clear that further fine-tuning may
improve performance and uncover further patterns.  Although automatic parameter
tuning may likely provide even more fine-tuned settings, our goal here is to
understand how UMM performs in comparison to an existing Bayesian method (CEGO)
and automatically tuning the parameters of CEGO is not computationally
feasible.  For consistency, in the following sections, we run UMM with this
parameter configuration. Moreover, we recommend this configuration as the
current default of UMM.

@@

To better understand the behaviour of the algorithms, we record the fitness of
the solution evaluated at each step of each run and we average those values
over 10 runs with different random seeds on the same problem instance.  As an example, each line in Fig.~\ref{fig:lop_synth} 
shows the mean fitness of the solutions evaluated by either UMM or CEGO on one synthetic LOP
instance. The shaded area shows one standard deviation around the
mean. Ideally, each new evaluation will monotonically improve in fitness, since each solution evaluated helps to refine the model and leads to a
better solution being evaluated in the next step.  However, it is often the case that  \EKHINE{yo diría solo que lo esperable es que baje monotonically, sin entrar a valorar un ejemplo concreto}\MANUEL{Así está mejor? Quiero llamar la atención porque normalmente se muestra la evolución de la mejor solución encontrada pero nosotros estamos mostrando el fitness de cada solución evaluada}
% the fitness of the solutions does not improve monotonically and
new solutions
are often worse than their predecessors. Nevertheless, in this example, there
is a clear overall improvement as more evaluations are performed. The minimum
values achieved in the plot up to a particular number of function evaluations
gives an estimation of the mean fitness of the best-so-far solution up to that
point in the run. The algorithms do not return the final solution shown but
rather the best one (minimum fitness) of each run.

@@

When we compare the behaviour of CEGO and UMM on synthetic LOP instances
(Fig.~\ref{fig:lop_synth}), we observe a clear pattern independently of the
uniformity of the instances. In particular, CEGO starts with a random sampling
of X solutions, with are quite poor as expected, followed by building and
updating the Kriging model based new evaluations, which leads to an impressive
improvement in less than 50 evaluations. However, progress is more or less
halted at this point and the model seems to have trouble generating better
solutions. Lower uniformity values (e.g., $\mu=5$) show worsening evaluations
despite the increase in solutions used to build the model. In summary, the
Kriging model clearly helps CEGO to quickly identify good solutions in very few
evaluations. However, for some reason, CEGO is not able to further improve
those solutions given more data. This may due to the model not being able to
produce accurate predictions or the underlying optimizer not being able to find
improved solutions using the model.\MANUEL{perhaps we should measure the
  prediction error and see if it grows or decreases. That would answer this
  question.}

On the other hand, UMM converges much more slowly than CEGO, only reaching the
same fitness values around evaluation 250. Yet, UMM is able to keep the rate of
improvement and completely overtake CEGO around evaluation 350. As in the case
of CEGO, lower uniformity values lead to worse performance, although the
difference is much smaller for UMM than it was for CEGO.\MANUEL{If there is anything else we can say here, please go ahead and say it}

@@

Results are quite different on the 
%The fast initial convergence of CEGO for synthetic LOP instances does not seem
%to translate to the
instances available in LOPLIB~\citep{}. We show in
Fig.~\ref{fig:loplib} three examples with different permutation size, however,
the results are consistent for other LOPLIB instances (complete results
available as supplementary material \supplement). Due to extremely long
runtimes of CEGO and constraints in our computing system, we set a maximum
wall-clock time of 3 days for each run of CEGO. On these instances, the
behaviour of CEGO and UMM is much more similar (at least up to the termination
point of CEGO). It appears that our synthetic LOP instances have a fitness
landscape that is very amenable to CEGO's Kriging model, whereas the LOPLIB
instances do not present such landscape. Comparing CEGO and UMM results, the
differences get smaller, in favour of UMM, with larger instance size. Our
conjecture is that building an accurate Bayesian model and searching for good
solutions on it becomes harder for larger fitness landscapes.
\EKHINE{Parece que el paper va del CEGO en vez del UMM :p. Yo diría que tanto UMM como CEGO encuentran soluciones que mejoran según avanzan, pero con diferencias de comportamiento: (1) las los optimos de UMM son '3 veces mejor', y esto yo creo que es importante, aunque en gran parte se debe a que (2) UMM hace 400 ejecuciones en 3, 6 y 11 horas y CEGO entre 50 y 130 en 5 dias (3) cego empeora segun n aumenta (y n=200 tampoco es enorme)}
\MANUEL{Eso es porque puedo interpretar los resultados de CEGO mejor que los de UMM. Si crees que hay alguna interpretación o conjetura sobre UMM que podemos mencionar aquí, sería bueno mencionarla.}
\MANUEL{Prefiero dejar el comentario sobre las mejores soluciones encontradas y el tiempo requerido para cuando hable de la tabla, así se puede ver todo junto}

@@

Table~\ref{tab:results} shows overall results over all instances evaluated (excluding the synthetic LOP instances). In particular, the mean fitness (and standard deviation) of the best solution found at the end of each run,
In terms of best fitness achieved on each run, and taking into account the
random behaviour observed in some instances and discussed above,

@@

In this paper, we have introduced UMM, a population-based probabilistic
algorithm based on an unbalanced Mallows model. The algorithm is designed for
black-box combinatorial problems on permutation landscapes and when the budget
of fitness evaluations is severely limited (up to $1000$ or less).  UMM is
specially well-suited for budget-limited scenarios that are still
time-sensitive, e.g., when the overhead incurred by the optimizer should be no
more than a few minutes added to each fitness evaluation. As shown by our
results, a Bayesian optimizer such as CEGO may sometimes require more than hour
per fitness evaluation. The time required to train the surrogate-model of CEGO
also sharply increases with permutation length, hence, UMM would be a
computationally feasible alternative for relatively large problem sizes.

Although we were inspired by previous work on ACO for expensive black-box
combinatorial problems~\citep{PerLopStu2015si}, UMM is, to the best of our
knowledge, the first EDA specifically designed for such problems that is able
to obtain competitive results when compared with Bayesian optimizers. The
results presented here show that, despite the intrinsic difficulty of this scenario, there are still significant advances yet to be made. Thus, we hope
that this paper will motivate further research. % on such problems.

A somewhat surprising result was that the behavior of UMM (and the same for
CEGO) on the PFSP is close to random. This result may be due to the distance
metric used, but further analysis is needed to corroborate this
conjecture. Here, we have focused on Kendall's $\tau$ distance, however, it is
possible to extend UMM to other distance metrics, which will allow us to
dynamically select among various distance metrics for an unknown black-box
permutation landscape~\citep{ZaeStoBar2014:ppsn}. Also, as shown by our
experiments, the behavior of the parameters ($r_1$ and $r_2$) of UMM is not
always obvious, thus a more detailed analysis would be needed to provide either
generally good static values or an online adaptation approach. Finally, we have
used here three very different combinatorial problems (LOP, PFSP and QAP) as
black-box benchmarks. However, an even more diverse range of problems would be
needed to understand the behavior of UMM on real-world black-box combinatorial
landscapes.

>>>>

\subsection{Permutations}

@@

The exact MLE for $\sigma_0$ is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}. Unfortunately,  obtaining such ranking is NP-hard~\cite{DwoKumNao2001rank}. An alternative to the Kemeny ranking is the Borda ranking~\cite{AliMei2011kemeny}, which has several advantages: (i) it requires polynomial computational time; (ii) it  guaranties  high quality parameters for a sample distributed according to the MM~\cite{CarProSha2013votes}; and (iii) and also when no distribution is assumed, Borda is an approximation to Kemeny~\cite{CopFleRur2010ordering}.

\subsection{Unbalanced Borda}\label{sec:uborda}
We introduce now the core for our probabilistic algorithm, the uBorda algorithm. Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking. The uBorda algorithm is a recent generalization of Borda~\cite{???}.\MANUEL{X This paper is missing \ekhine{}{to be pocessed by arxiv}} In uBorda, we consider that each permutation in the sample $\sigma\in S$ has an associated weight $w(\sigma)$ where $w\colon S_n \to \mathbb{R^{+}}$. Then, the score of \manuel{item}{position} $i$ is defined as follows:
%
\begin{equation}\label{eq:uborda_score}
B(i) =  \sum_{t\in S}  w(\sigma_t) \sigma_t(i) \enspace,
\end{equation}
%
and, the same as Borda, orders the \ekhine{items}{positions}\MANUEL{X why items? we talk about positions above. We could talk about decision variables if that is clearer\ekhine{}{id rather use the simplest possible notation. In statistical literature you ll see items/rankings. }} increasingly by their score.

@@

@@

As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $\minit$ permutations that are uniformly randomly generated (although it would be possible to employ Latin Hyper-Cube sampling or other methods) and evaluated. Then, using the current sample $S$, the uBorda algorithm is applied to find a reference permutation $\hat{\sigma}_0$ (line~\ref{line:uborda}). 
Together with the parameter $\theta$, they define the Mallows model $MM(\hat\sigma_0, \theta)$. 
A decreasing $\theta$ parameter (line~\ref{line:theta}) implies a decreasing variance of the Mallows model (Eq.~\ref{eq:MM}), as explained in  Section~\ref{sec:sampling} below. 
Then, we sample a new candidate permutation from  $MM(\hat\sigma_0, \theta)$ (line~\ref{line:sample}). This candidate permutation (solution) is evaluated and added to the sample $S$. The process is repeated up to a given maximum budget of evaluations $\FEmax$.\MANUEL{If the learning step happens later, when does it happen here?}
\MANUEL{We do not explain the initial value of $\theta$ or how we decrease it. Can we give a bit more detail?\ekhine{no without explaining the expecteattions, too messy}}

@@

@@

\subsubsection{Learning step.}
At each iteration we start by fitting a MM to the sample $S$, which implies
estimating $\hat\sigma_0$ and $\theta$. Parameter $\theta$ is updated
deterministically at the sampling step.  The parameter $\sigma_0$, however, is
learned from the current sample $S$ by using the uBorda algorithm with a weight
$w(\sigma)=\rho^{f(\sigma)}$ (Section~\ref{sec:uborda}). The learning rate is
controlled by parameter $\rho \in (0,1)$. %
In UMM, the value of $\rho$ depends on the set of true fitness evaluations $F$
and it is set at each iteration such that the largest $100r_1$\% of the mass of
the weights is concentrated in the best $100r_2$\% of the solutions in $S$, i.e.,
\MANUEL{Let's talk about this Equation tomorrow}
%
\begin{equation}
\begin{split}
r_1 \sum_{\sigma\in S}\rho^{f(\sigma)} =  \sum_{\sigma\in S'}\rho^{f(\sigma)} \quad\text{where}\quad |S'| = r_2|S| \land \forall \sigma'\in S', \sigma\in S\setminus S' : f(\sigma')\geq f(\sigma)
\end{split}
\end{equation}
%
\MANUEL{X should the $\sigma$ be the same on both sides of $=$?}
where $S'$ contains the $r_2|S|$ elements of $S$ with the best corresponding fitness values in $F$. The values of $r_1, r_2 \in (0, 1)$ are user-defined and control the balance between exploration and intensification: higher values bias the sampling towards the best solutions found so far, which may lead to quick stagnation, whereas smaller values make the sampling more uniform, thus slowing down the learning.

@@

\subsubsection{Sampling step.}\label{sec:sampling}
Sampling from $MM(\sigma_0,\theta)$  can be done efficiently~\cite{IruCalLoz2016permallows} given the parameters  $\sigma_0$ and $\theta$, as studied by \citet{Collas,IruCalLoz2016permallows,DoiPekReg2004rank}\MANUEL{Which paper is Collas? It is not found.} where the computational complexities are $O(n\log n)$ and $O(n^2)$.\MANUEL{X respectively? more than 2}
%
At each iteration of UMM, $\sigma_0=\hat\sigma_0$ is the permutation obtained in the previous learning step. The scale $\theta$ controls the variance and expected distance $\Edist$ of the MM, i.e., $D$ is the distance of a random permutation $\sigma\sim MM(\sigma_0, \theta)$ to the location parameter $\sigma_0$~\cite{gMallows}:\MANUEL{X where does this formula come from?\ekhine{}{}}
%
\begin{equation}\label{eq:expectation}
\Edist = \frac{n \cdot \exp(-\theta)}{1-\exp(-\theta)} - \sum_{j=1}^{n}  \frac{j\cdot\exp(-j  \theta)}{1-\exp(-j\theta)}
\end{equation}

@@

A purely explotation approach would always generate $\sigma_0$ in the sampling stage. However, the setting of a decreasing variance \MANUEL{But we are controlling the expectation not the variance???\ekhine{}{both are a function of $\theta$}}  controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.  Similar approaches have been taken in different contexts~\cite{ArzCebPer2019qap}.

@@

\subsubsection{Other}.  For simplicity, we use Kendall's $\tau$
distance~(Eq.~\ref{eq:kendall_decomp}) in all
experiments\MANUEL{\citet{ZaeStoBar2014:ppsn} claims this is called Swap
  distance, could you double-check?\ekhine{}{es la kendall among the inverses en realidad}}. Nevertheless,
\citet{ZaeStoFriFisNauBar2014} points out that CEGO performs best on the QAP
when using Hamming distance.  We plan to extend UMM to other distance measures
and to the dynamic selection of distance measures~\citep{ZaeStoBar2014:ppsn}.
In all experiments, we consider a maximum budget of $\FEmax=400$ evaluations of
the actual objective function. In a white-box context, state-of-the-art
algorithms for the problems considered here typically evaluate
thousands of solutions, thus, the budget considered
here for the black-box context is extremely limited.

@@

\subsubsection{Computing environment.}
All experiments were run on Intel Xeon ``Ivybridge'' E5-2650v2 CPUs at
2.60\,GHz, 64\,GB RAM running CentOS Linux release 7.4.1708 (Core).
\MANUEL{This may need updating since you ran some experiments? didn't you?}
\ekhine{}{esto es lo q se: Hipatia (Cluster)
18 nodes (1 with Nvidia Tesla K40 GPU)
624 cores
4TB of RAM Memory
Infiniband network}

@@

\subsection{Setting the learning rate}
The first part of the experimental analysis concerns the {analyses} of the learning rate, which corresponds to parameter $\rho$ of uBorda. As stated in Section~\ref{sec:sampling}, $\rho$ is automatically set at each iteration in such a way that the $100r_1\%$ of the samples have the $100r_2\%$ of the weight. In this regard, we tried different configurations of parameters $r_1$ and $r_2$ in all the instances.\MANUEL{?? why did we try this values.} Figure~\ref{fig:heatmaps} shows the best solution\MANUEL{best solution found at the end of a run?} (averaged for 10 runs) for instances: \texttt{N-t1d100.01} and \texttt{N-t1d150.01} for the LOP, \texttt{reC05} for the PFSP, \texttt{kra32} for the QAP. This experiment for all the real datasets can be bound in the supplemental material. \MANUEL{The caption says synthetic instances}

\begin{figure}[htb]
  \centering%
  \includegraphics[width=0.24\textwidth]{../img/heatmap_N-t1d100_01}
  \includegraphics[width=0.24\textwidth]{../img/heatmap_N-t1d150_01}
  \includegraphics[width=0.24\textwidth]{../img/heatmap_rec05}
  \includegraphics[width=0.24\textwidth]{../img/heatmap_kra32_dat}
    \caption{Mean fitness  for different configuration for the learning rate \label{fig:heatmaps}}
\end{figure}

@@

Over all the instances \MANUEL{X (which instances we tested? all synthetic?)} we
see that the configuration $r_1 =0.1$, $r_2=0.9$ has good
performance\MANUEL{how do you define this? The best mean?}. From this
preliminary experiment, it is not obvious how the best settings are related to
problem features, such as permutation length ($n$). On the other hand, given
the sharp transitions in the plots, it is clear that further fine-tuning may
improve performance and uncover further patterns.  Although automatic parameter
tuning may likely provide even more fine-tuned settings, our goal here is to
understand how UMM performs in comparison to an existing Bayesian method (CEGO)
and automatically tuning the parameters of CEGO is not computationally
feasible.  For consistency, in the following sections, we run UMM with this
parameter configuration. Moreover, we recommend this configuration as the
current default of UMM.

@@

To better understand the behaviour of the algorithms, we record the fitness of
the solution evaluated at each step of each run and we average those values
over 10 runs with different random seeds on the same problem instance.  As an example, each line in Fig.~\ref{fig:lop_synth} 
shows the mean fitness of the solutions evaluated by either UMM or CEGO on one synthetic LOP
instance. The shaded area shows one standard deviation around the
mean. Ideally, each new evaluation will monotonically improve in fitness, since each solution evaluated helps to refine the model and leads to a
better solution being evaluated in the next step.  However, it is often the case that 
new solutions
are often worse than their predecessors. Nevertheless, in this example, there
is a clear overall improvement as more evaluations are performed. The minimum
values achieved in the plot up to a particular number of function evaluations
gives an estimation of the mean fitness of the best-so-far solution up to that
point in the run. The algorithms do not return the final solution shown but
rather the best one (minimum fitness) of each run.

@@

\EKHINE{cambio un poco al argumento}
When we compare the behaviour of CEGO and UMM on synthetic LOP instances
(Fig.~\ref{fig:lop_synth}), we observe a clear pattern independently of the
uniformity of the instances. In particular, we see a linear convergence to the optimal value for UMM. This is due to the linear decrease in the parameter $\theta$ (line~\ref{line:theta} in Algorithm~\ref{alg:umm}), which imposes a linear variance decrease on the internal probability model of the algorithm over the $m$ functions evaluations. Overall, UMM found the best solution in all the repetitions.

%UMM converges much more slowly than CEGO, only reaching the
%same fitness values around evaluation 250.

%Yet, UMM is able to keep the rate of
%improvement and completely overtake CEGO around evaluation 350. As in the case
%of CEGO, lower uniformity values lead to worse performance, although the
%difference is much smaller for UMM than it was for CEGO.\MANUEL{If there is anything else we can say here, please go ahead and say it}

On the other hand, \EKHINE{cambio el orden}
CEGO starts with a random sampling
of X\ekhine{}{10?} solutions, with are quite poor as expected, followed by building and
updating the Kriging model based new evaluations, which leads to an impressive
improvement in less than 50 evaluations. However, progress is %more or less 
halted at this point and the model seems to have trouble generating better
solutions. Lower uniformity values (e.g., $\phi=.5$) show worsening evaluations
despite the increase in solutions used to build the model. In summary, the
Kriging model clearly helps CEGO to quickly identify good solutions in very few
evaluations. However, for some reason, CEGO is not able to further improve
those solutions given more data. This may due to the model not being able to
produce accurate predictions or the underlying optimizer not being able to find
improved solutions using the model.\MANUEL{perhaps we should measure the
  prediction error and see if it grows or decreases. That would answer this
  question.}

@@

%Results are quite different on the 
%The fast initial convergence of CEGO for synthetic LOP instances does not seem
%to translate to the
%instances available in LOPLIB~\citep{}. We show in
%Fig.~\ref{fig:loplib} three examples with different permutation size, however,
%the results are consistent for other LOPLIB instances (complete results
%available as supplementary material \supplement). Due to extremely long
%runtimes of CEGO and constraints in our computing system, we set a maximum
%wall-clock time of 3 days for each run of CEGO. On these instances, the
%behaviour of CEGO and UMM is much more similar (at least up to the termination
%point of CEGO). It appears that our synthetic LOP instances have a fitness
%landscape that is very amenable to CEGO's Kriging model, whereas the LOPLIB
%instances do not present such landscape. Comparing CEGO and UMM results, the
%differences get smaller, in favour of UMM, with larger instance size. Our
%conjecture is that building an accurate Bayesian model and searching for good
%solutions on it becomes harder for larger fitness landscapes.
%\EKHINE{Parece que el paper va del CEGO en vez del UMM :p. Yo diría que tanto UMM como CEGO encuentran soluciones que mejoran según avanzan, pero con diferencias de comportamiento: (1) las los optimos de UMM son '3 veces mejor', y esto yo creo que es importante, aunque en gran parte se debe a que (2) UMM hace 400 ejecuciones en 3, 6 y 11 horas y CEGO entre 50 y 130 en 5 dias (3) cego empeora segun n aumenta (y n=200 tampoco es enorme)}
%\MANUEL{Eso es porque puedo interpretar los resultados de CEGO mejor que los de UMM. Si crees que hay alguna interpretación o conjetura sobre UMM que podemos mencionar aquí, sería bueno mencionarla.}
%\MANUEL{Prefiero dejar el comentario sobre las mejores soluciones encontradas y el tiempo requerido para cuando hable de la tabla, así se puede ver todo junto}

In the following paragraphs we analyze the result over real rata. Fig.~\ref{fig:loplib} shows the results on the instances available in LOPLIB~\citep{}. The distribution of this instances differs from the symthetic LOP instances, and thus so does the landscape. However, UMM shows a very similar performance: there is a (close to) linear decrease on the function evaluations of new solutions along iterations. This suggests that UMM very good fits the fitness landscape. Regarding CEGO, we can take similar conclussions for the first 50-100 iterations. However, the processes had to be halted after hitting the wall time, so no conlusion can be taken for further iterations. Overall, the improvement from the first (random) to the last solutions is 2-3 times larger in UMM than CEGO.

@@

Table~\ref{tab:results} shows overall results over all instances evaluated (excluding the synthetic LOP instances). In particular, the mean fitness (and standard deviation) of the best solution found at the end of each run,
In terms of best fitness achieved on each run, and taking into account the
random behaviour observed in some instances and discussed above,\ekhine{}{falta un cacho aqui}

@@

In this paper, we have introduced UMM, a population-based probabilistic
algorithm based on an unbalanced Mallows model. The algorithm is designed for
black-box combinatorial problems on permutation landscapes and when the budget
of fitness evaluations is severely limited (up to $1000$ or less).  UMM is
specially well-suited for budget-limited scenarios that are still
time-sensitive, e.g., when the overhead incurred by the optimizer should be no
more than a few minutes added to each fitness evaluation. As shown by our
results, a Bayesian optimizer such as CEGO may sometimes require more than hour
\EKHINE{me parece mas de CEGO q UMM y parece q UMM solo tiene de ventaja ir rápido. Primero destacaría lo bueno de UMM. Creo q es importante destaacar q UMM va MEJOR o IGUAL en TODOS los experimentos  y luego que addemás es más rápido}
per fitness evaluation. The time required to train the surrogate-model of CEGO
also sharply increases with permutation length, hence, UMM would be a
computationally feasible alternative for relatively large problem sizes.

Although we were inspired by previous work on ACO for expensive black-box
combinatorial problems~\citep{PerLopStu2015si}, UMM is, to the best of our
knowledge, the first EDA specifically designed for such problems that is able
to obtain competitive\EKHINE{hhmmm lo hacemos mejor o igual} results when compared with Bayesian optimizers. The
results presented here show that, despite the intrinsic difficulty of this scenario, there are still significant advances yet to be made. Thus, we hope
that this paper will motivate further research. % on such problems.

A somewhat surprising result was that the behavior of UMM (and the same for
CEGO) on the PFSP is close to random. This result may be due to the distance
metric used, but further analysis is needed to corroborate this
conjecture. Here, we have focused on Kendall's $\tau$ distance, however, it is
possible to extend UMM to other distance metrics, which will allow us to
dynamically select among various distance metrics for an unknown black-box \ekhine{}{estoy totalmente en contra de esto :) }
permutation landscape~\citep{ZaeStoBar2014:ppsn}. Also, as shown by our
experiments, the behavior of the parameters ($r_1$ and $r_2$) of UMM is not
always obvious, thus a more detailed analysis would be needed to provide either
generally good static values or an online adaptation approach. Finally, we have
used here three very different combinatorial problems (LOP, PFSP and QAP) as
black-box benchmarks. However, an even more diverse range of problems would be
needed to understand the behavior of UMM on real-world black-box combinatorial
landscapes.
